% ChemmineRdata
% Kevin Horan & Thomas Girke
% `r format(Sys.time(),"%B %d, %Y")`
<!--
\usepackage[utf8]{inputenc}
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{ChemmineRdata}
-->

```{r echo=FALSE, results='hide',message=FALSE,warning=FALSE,error=FALSE}
library(ChemmineRdata) 
library(knitcitations)
library(RSQLite)
```


Introduction
============
This package contains two datasets, Drug Bank and DUD. They are stored in 
an SQLite database along with several indexed features derived from each 
compound. This allows one to quickly search for compounds by any of these features,
or just to see the value of the features without having to compute them each
time.  These datasets can be used with tthe tools provided by *ChemmineR*.

The DUD dataset was downloaded from [here](http://dud.docking.org/jahn/DUD\_LIB\_VS\_1.0.tar.gz). A description
can be found [here](http://dud.docking.org/jahn/Readme.pdf). 

The Drug Bank data set is version 4.1. It can be downloaded [here](http://www.drugbank.ca/system/downloads/current/structures/all.sdf.zip)


Features
========
The following features are included:

 - **aromatic**: Number of aromatic rings
 - **cansmi**: Canonical SMILES sting
 - **cansmins**:
 - **formula**: Molecular formula
 - **hba1**:
 - **hba2**:
 - **hbd**:
 - **inchi**: INCHI string
 - **logp**:
 - **mr**:
 - **mw**: Molecular weight
 - **ncharges**:
 - **nf**:
 - **r2nh**:
 - **r3n**:
 - **rcch**:
 - **rcho**:
 - **rcn**:
 - **rcooh**:
 - **rcoor**:
 - **rcor**:
 - **rings**:
 - **rnh2**:
 - **roh**:
 - **ropo3**:
 - **ror**:
 - **title**:
 - **tpsa**:

The DUD database additionally includes:

 - **target_name**: Name of the target
 - **type**: either "active" or "decoy"

Searching
=========

Search is done with the *findCompounds* function from the *ChemmineR* package. 
This function takes a connection to the desired database, a list of features you intend
to use in the tests and then a list of tests to use. It returns a list of compound id numbers. These are not
related to the names of the compounds in any way however, they are relevent only in the context of this database 
and may even change from one release to another. 

First create a connection to the desired database:

```{r }
dudConn = DUD()
```

The tests parameter is just a list of expressions like "FEATURE-NAME OPERATOR VALUE". For example,
if you want to find things with molecular weight less than 200 your test would look like "mw < 200".
For those who know SQL, any valid SQL expression can be used here. 


```{r }
results = findCompounds(dudConn,"mw",c("mw < 200"))
message("found ",length(results))
```

If more than one test is given, only compounds which satisify all tests are found. So if we wanted
to further restrict our search to compounds with 2 or more aromatic rings we could do:
```{r }
results = findCompounds(dudConn,c("mw","aromatic"),c("mw < 200","aromatic >= 2"))
message("found ",length(results))
```

Remember that any feature used in some test must be listed in the second argument.

String patterns can also be used. So if we wanted to match a substring of the molecular formula, so 
to find compounds with 6 carbon atoms, we could do:
```{r }
results = findCompounds(dudConn,"formula",c("formula like '%C6%'"))
message("found ",length(results))
```

The "like" operator does a pattern match. There are two wildcard 
operators that can be used with this operator. The "%" will match any stretch of characters while the "?"
wil match any single character. So the above expression would match a formula like "C6H12O6".

Valid comparison operators are: 

 - <,    <=,   > ,   >=
 - =,    ==,   !=,   <>,   IS,   IS NOT,   IN,   LIKE 

The boolean operators "AND" and "OR" can also be used to create more complex expressions within a single test.


Using Search Results
=====================
Once you have a list of compound ids from the *findCompounds* function, you can either
fetch the compound names, or the whole set of compounds as an SDFset. 

```{r }

#get the names of the compounds:
names = getCompoundNames(dudConn,results)

#if the name order is important set keepOrder=TRUE 
#It will take a little longer though
names = getCompoundNames(dudConn,results,keepOrder=TRUE) 


# get the whole set of compounds
sdfset = getCompounds(dudConn,results)
#in order:
sdfset = getCompounds(dudConn,results,keepOrder=TRUE)
#write results directly to a file:
sdfset = getCompounds(dudConn,results,filename=file.path(tempdir(),"results.sdf"))
```




```{r biblio, echo=FALSE, results='asis'}
#   bibliography()
```
